\bf Mainstreaming Race Science

### 2023-06-29 paper/paper.Rmd
working on paper

@@

We focus here on three topics identified in one of the 24 fitted topic models *[composite fig: topics over time + Silge plots]*.  In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.  The Silge plot contains racial terms (`races`, `whites` and potentially `europe`, `africa`, and `india`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.

*[switch 05 with 22; or add together 05, 22, 35, 38?]*
Topic 05 is strongly associated with I and PID in the same way.  The Silge plot contains the names of two Pioneer-funded researchers, `jensen` and `vernon`, but no racial terms.  (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model. Topic 22 )  The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.  Qualitative review of the top 121 papers in this topic (those with $gamma > 0.97$) confirmed that this topic has high precision for race science discourse, with *[x% (n/121)]* classified as race science discourse by the authors.

In a given year, most papers in topic 24 are published in mainstream journals, not MQ.  Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} is published in 1969 in Harvard Educational Review (not included in this study), and the time series suggests an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).  MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, *[confirming the claim that Lynn shifted MQ's focus to intelligence research]*.  PID publishes multiple papers in topic 24 almost immediately after it is founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.

Notably, BG published very few articles in this topic.  Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public *[cites/ev]*.  Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.  In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.  Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24.

@@

The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had begun to distance itself from race science.  *[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*

@@

@@

We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.  This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.  This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.  *[availability]*

@@

### Topic quality assessment

>>>>

*[write this]*

@@

![Silge plots and smoothed time series for three focal topics.  Top row: top 15 terms by $\beta$ (term-topic distribution) for each topic.  Bottom row: count of articles associated with each topic, by journal and year.  Article counts use a threshold approach, with $\gamma$ (topic-document distribution) greater than 0.5.  Thin lines give annual values, thick lines give 5-year running averages.  Because *Behavior Genetics* and *Behavioral and Brain Sciences* are not prominent in any panel, neither is given a direct label.](img/focal_topics.png){width="4.76in" height="3.57in"}

We focus here on three topics identified in one of the 24 fitted topic models *[composite fig: topics over time + Silge plots]*.  In this model, topic 07 is strongly associated with MQ:  MQ published dozens of articles in this topic each year, and no other journal ever published more than a handful, with most other journals publishing none in a given year.  The Silge plot contains racial terms (`races`, `whites`, `negroes` and potentially `europe`, `africa`, `india`, and `japan`) as well as `book`, likely reflecting the fact that MQ published a number of book reviews while the psychology journals did not.

Topic 22 is strongly associated with I and PID in the same way.  The Silge plot does contain `jensen`, as well as a reference to Raymond Cattell, who played a major role in the development of factor analysis and intelligence testing also advocated eugenics, fascism, and Nazi race science\cite{MehlerBeyondismRaymondCattell1997}. However, the other authors named in this topic — John Horn and Peter Bentler — do not appear to have contributed to race science discourse.  Instead this topic appears to identify "mainstream" (non-race science) intelligence research, specifically on factor analysis and the debate over whether intelligence in unidimensional or multidimensional.  (Three other topics only associated with mainstream intelligence journals and terms were also identified by this topic model.)  This topic indicates that the model is not simply lumping race science research together with other intelligence research.

The Silge plot for topic 24, by contrast, suggests a topic focused squarely on race and intelligence, with multiple racial terms (`whites`, `blacks`, `racial_differences`, `races`, `race_differences`) and the names of three prominent Pioneer-funded researchers, `jensen`, `lynn`, and `rushton`.  To confirm this impression from the Silge plot, we independently coded the top 121 papers in this topic (those with $gamma > 0.97$) as contributing to race science discourse or not.  105 (87%) were coded as race science discourse by both authors; 10 (8.3%) were coded as not race science discourse by both; and discordant codes were given to the remaining 6 (5.0%) (95% agreement, Cohen's $\kappa = 0.74$).

In almost all years, most papers in topic 24 were published in mainstream journals rather than MQ.  Jensen's "How Much Can We Boost IQ and Scholastic Achievement?" \cite{JensenHowMuchCan1969} was published in 1969 in Harvard Educational Review (not included in this study), and the time series indicates that, around this time, there was an increase in articles in topic 24 in both PS and MQ (the only two journals in our corpus that were active at the time).  MQ shows another sharp increase in the late 1970s, around the time Lynn was brought on as assistant editor, *[confirming the claim that Lynn shifted MQ's focus to intelligence research]*.  PID published multiple papers in topic 24 almost immediately after it was founded, with I showing a more gradual increase between the mid-1970s and mid-1990s.

Notably, BG published very few articles in this topic.  Outside of the field itself, behavior genetics is strongly associated with race science by both other academics and the general public.  Panofsky \cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to Jensen's 1969 paper, behavior genetics emphasized the study of non-human animals and intentionally avoided associations with eugenics and public controversy more generally.  In response to Jensen, critics such as Lewontin offered broad critiques of behavior genetics as such, and behavior geneticists in turn adopted a radical conception of academic freedom (without any sense of responsibility for the social implications of academic research) and a siege or wartime mentality, as demonstrated by Sandra Scarr's 1986 presidential address to the Behavior Genetics Association (BGA) \cite{ScarrThreeCheersBehavior1987}.  Scarr's address coincided with the period between 1970-1990 when BG published articles in topic 24 (including Scarr's address itself).

@@

The topic model analysis and the case of Whitney's address suggest that, by the 1990s, behavior genetics had at least somewhat distanced itself from race science.  *[https://link.springer.com/search?use-oscar-shared-search=true&sortBy=newestFirst&dateTo=2010&sortOrder=newestFirst&facet-journal-id=10519&content-type=Article&query=race+intelligence&date=custom&dateFrom=&date-facet-mode=between&facet-start-year=1972&previous-start-year=1972&facet-end-year=2010&previous-end-year=2023]*

@@

# Discussion

Our findings indicate that *Mankind Quarterly* (MQ) was less important as an echo chamber for late 20th century race science than mainstream psychology journals, especially *Intelligence* (I) and *Personality and Individual Differences* (PID).  Indeed, MQ published a very different kind of race science from the race-and-intelligence research published in I and PID, even after *[Lynn joined the MQ editorial board]*.

We stress that we are not identifying any particular scientist as a racist based on the topic model results.  Scientific research can be utilized to promote scientific racism — and thus count as race science — independently of the views and attitudes of the scientist who conducted that research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}  But this does not mean that scientists are not responsible for the downstream social effects of their 
research, even when these effects involve deliberate misinterpretation.*[Douglas, Kitcher, Kourany, Havstad; Block and Dworkin?]*  Like other citizens, scientists are responsible for mitigating the reasonably foreseen harms that result from their actions.  And we now have decades of empirical evidence that research on the genetics of intelligence — even research that does not itself touch on race — can and will be used to promote scientific racism.\cite{MeyerWrestlingSocialBehavioral2023}

*[engagement dilemma: technical critiques can legitimize]*

*[this probably goes too far even for a discussion section]*

Proponents of a strong conception of academic freedom often point to chapter 2 of philosopher John Stuart Mill's *On Liberty* \cite{MillLiberty1862}, "Of the Liberty of Thought and Discussion."  However, in chapter 1 of the same book, Mill states the Harm Principle as the overall thesis of his book: "The only end for which people are entitled, individually or collectively, to interfere with the liberty of action of any of their number is self-protection. The only purpose for which power can be rightfully exercised over any member of a civilized community, against his will, is to prevent harm to others." Even for Mill, free speech and free inquiry are appropriately restrained when the ideas in question will harm others.

More recently, philosopher Shannon Dea \cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is importantly distinct from freedom of speech.  Whereas freedom of speech is universal, enjoyed by all residents of countries where speech is legally protected, academic freedom is more exclusive, enjoyed only by college and university faculty.  (Though, as Dea points out, non-tenure-track faculty often do not have effective protections for academic freedom.)  Freedom of speech primarily serves to protect the interests of the individual speaker: that you will not be punished by the state for stating unpopular opinions.  But academic freedom is "other-directed," primarily serving to benefit the common good of society as a whole by enabling the pursuit of truth.

Advocates of research on genetics and intelligence — both anti-racist liberals and race scientists — claim this work will lead to social benefits.  This reflects a recognition that the purpose of academic freedom is to promote the common good, rather than merely satisfy an individual's curiosity.  But the promised benefits are either vague or do not actually require understanding the genetics of intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}

Using agent-based models, Weatherall et al.\cite{WeatherallHowBeatScience2018} show that a strategy of "selective sharing" — promoting favorable results from independent research — can be effective at misleading a target audience.

@@

*[drop]*
We also adapted the "discursive space" analysis proposed by *[Hicks 2021?]*.  This analysis combines the Hellinger distance between topic-document distributions — capturing how "far apart" two documents are — with dimensional reduction techniques used for data visualization.  This produces a two-dimensional visualization of individual documents, arranged so that documents with similar topic distributions are visually clustered and visually distinct from documents with different topic distributions.  *[availability]*

@@

### Topic quality assessment
*[rewrite]*

### 2023-06-29 paper/paper.tex
working on paper

\begin{sciabstract}
Enter the text of your abstract here.
\end{sciabstract}

@@

@@

@@

\begin{longtable}[]{@{}lrl@{}}
\caption{Pioneer-funded researchers. Either identified in
\emph{{[}cite{]}} or named on Pioneer's website, along with identified
discipline and WoS author search result counts.}\tabularnewline
\toprule
\endhead
Thomas J. Bouchard, Jr. & psychology & \\
Brunetto Chiarelli & anthropology? & \\
Hans Eysenck & psychology & \\
Robert Gordon & sociology & \\
Linda Gottfredson & psychology & \\
Garrett Hardin & ecology & \\
Joseph M. Horn & psychology & \\
Lloyd Humphreys & psychology & \\
Arthur Jensen & psychology & \\
Michael Levin & philosophy & \\
Richard Lynn & psychology & \\
R. Travis Osborne & psychology & \\
J. Phillippe Rushton & psychology & \\
Audrey M. Shuey & psychology & \\
Philip A. Vernon & psychology & \\
Daniel Vining, Jr. & demography & \\
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
\caption{Journals publishing 6 or more Pioneer-funded researchers, WoS
author search results \emph{{[}date coverage{]}}}
\end{figure}

Figure \emph{{[}fig{]}} shows that, while MQ is on this list of
``Pioneer-publishing'' journals, a number of mainstream journals are
more prominent: \emph{Personality and Individual Differences} (PID),
\emph{Intelligence} (I), \emph{Behavior Genetics} (BG), and
\emph{Psychological Reports} (PR). In addition, only psychologist
Richard Lynn appears to have published heavily in MQ. Lynn has been an
assistant editor of MQ since 1979 \emph{{[}check this:
\url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president
of PF since the death of psychologist J. Phillippe Rushton in 2012
\emph{{[}cite{]}}. By contrast, a number of Pioneer-funded researchers
have published in PID, I, and to a lesser degree BG: Eysenck, Jensen,
Rushton, Vernon, and also Lynn.

PID was founded in 1980, with Eysenck as editor-in-chief and an
editorial board including Jensen and Lynn. In the inaugural editorial,
Eysenck identified ``studies of the genetic determinants of individual
differences in the areas of personality and intelligence'' as one of the
journal's eight major areas of interest. Eysenck remained
editor-in-chief until his death in 1997. In 2005 the editorial board
still included Jensen and Lynn. PID was first published by Pergamon
Press, a mainstream academic press, and today is published by Elsevier.
According to the Scimago Journal Rankings, based on Scopus citation
data, in 2022 PID had an H-index of 193, \#29 in Psychology, and a
citation index of 5.27, \#24 in Psychology.

\emph{{[}similar para on I{]}}

\emph{{[}segue to topic model analysis{]}}

@@

>>>>

\begin{sciabstract}
\emph{{[}structured{]}}
\end{sciabstract}

@@

\emph{{[}write this{]}}

@@

\emph{{[}for xref:
\url{https://bookdown.org/yihui/bookdown/a-single-document.html}{]}}

@@

\begin{longtable}[]{@{}lrl@{}}
\caption{Pioneer-funded researchers. Either identified in
\emph{{[}cite{]}} or named on Pioneer's website, along with identified
discipline and WoS author search result counts.}\tabularnewline
\toprule
\endhead
Thomas J. Bouchard, Jr. & psychology & \\
Brunetto Chiarelli & anthropology & \\
Hans Eysenck & psychology & \\
Robert Gordon & sociology & \\
Linda Gottfredson & psychology & \\
Garrett Hardin & ecology & \\
Joseph M. Horn & psychology & \\
Lloyd Humphreys & psychology & \\
Arthur Jensen & psychology & \\
Michael Levin & philosophy & \\
Richard Lynn & psychology & \\
R. Travis Osborne & psychology & \\
J. Phillippe Rushton & psychology & \\
Audrey M. Shuey & psychology & \\
Philip A. Vernon & psychology & \\
Daniel Vining, Jr. & demography & \\
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=4.76in,height=2.6in]{img/wos_results.png}
\caption{Journals publishing 6 or more Pioneer-funded researchers, Web
of Science author search results \emph{{[}date coverage{]}}}
\end{figure}

Figure \emph{{[}fig{]}} shows that, while MQ is among the
``Pioneer-publishing'' journals, a number of mainstream journals are
more prominent: \emph{Personality and Individual Differences} (PID),
\emph{Intelligence} (I), \emph{Behavior Genetics} (BG), and
\emph{Psychological Reports} (PR). In addition, only psychologist
Richard Lynn appears to have published heavily in MQ. Lynn has been an
assistant editor of MQ since 1979 \emph{{[}check this:
\url{https://rationalwiki.org/wiki/Mankind_Quarterly}{]}} and president
of PF since the death of psychologist J. Phillippe Rushton in 2012
\emph{{[}cite{]}}. By contrast, a number of Pioneer-funded researchers
have published in PID, I, and to a lesser degree BG: Bouchard, Eysenck,
Jensen, Rushton, Vernon, and also Lynn.

\emph{Personality and Individual Differences} (PID) was founded in 1980,
with Eysenck as editor-in-chief and an editorial board including Jensen
and Lynn. In the inaugural editorial, Eysenck identified ``studies of
the genetic determinants of individual differences in the areas of
personality and intelligence'' as one of the journal's eight major areas
of interest. Eysenck remained editor-in-chief until his death in 1997.
In 2005 the editorial board still included Jensen and Lynn. PID was
first published by Pergamon Press, a mainstream academic press, and
today is published by Elsevier. According to the Scimago Journal
Rankings, based on Scopus citation data, in 2022 PID had an H-index of
193, \#29 in Psychology (Miscellaneous), and a 2-year citation index
(impact factor) of 5.27, \#24 in Psychology (Miscellaneous).

\emph{Intelligence} (I) was founded in 1977, with psychologist Douglas
Detterman as editor-in-chief from the founding until 2016. Lloyd
Humphreys was on the editorial board starting from 1977; by 1990 he had
been joined by Jensen and Philip Vernon. Richard Lynn joined the
editorial board sometime between 1998 and 2002. (Archive copies of the I
editorial board page are not available from the journal's website from
1999 through 2001.) I has been criticized for including Lynn and Gerhard
Meisenberg on its editorial board until July-August 2018
\cite{SainiSuperiorReturnRace2019}. Meisenberg is also connected with
the Pioneer Fund \emph{{[}director?{]}}, and was editor-in-chief of MQ
in 2015-18 \emph{{[}check these{]}}. I is published by Elsevier.
According to the Scimago Journal Rankings, based on Scopus citation
data, in 2022 I had an H-index of 103, \#32 in Experimental and
Cognitive Psychology, and a 2-year citation index (impact factor) of
3.05, \#31 in Experimental and Cognitive Psychology.

@@

The fact that Pioneer-funded researchers published heavily in two
mainstream psychology journals does not tell us anything about the
content of their publications or the findings of their research. We
assembled a full-text corpus of articles published in MQ, PID, I, BG,
PR, and Behavior and Brain Sciences (BBS) between 1960-2010 and used
topic modeling to identify topics discussing race, intelligence, and
both. \emph{{[}tables: Ns for each article, topic model parameters;
figs: articles by pub over time, topic heatmaps, Silge plots,
race-intelligence-both grids{]}}

\begin{figure}
\centering
\includegraphics[width=4.76in,height=3.57in]{img/focal_topics.png}
\caption{Silge plots and smoothed time series for three focal topics.
Top row: top 15 terms by \(\beta\) (term-topic distribution) for each
topic. Bottom row: count of articles associated with each topic, by
journal and year. Article counts use a threshold approach, with
\(\gamma\) (topic-document distribution) greater than 0.5. Thin lines
give annual values, thick lines give 5-year running averages. Because
\emph{Behavior Genetics} and \emph{Behavioral and Brain Sciences} are
not prominent in any panel, neither is given a direct label.}
\end{figure}

We focus here on three topics identified in one of the 24 fitted topic
models \emph{{[}composite fig: topics over time + Silge plots{]}}. In
this model, topic 07 is strongly associated with MQ: MQ published dozens
of articles in this topic each year, and no other journal ever published
more than a handful, with most other journals publishing none in a given
year. The Silge plot contains racial terms (\texttt{races},
\texttt{whites}, \texttt{negroes} and potentially \texttt{europe},
\texttt{africa}, \texttt{india}, and \texttt{japan}) as well as
\texttt{book}, likely reflecting the fact that MQ published a number of
book reviews while the psychology journals did not.

Topic 22 is strongly associated with I and PID in the same way. The
Silge plot does contain \texttt{jensen}, as well as a reference to
Raymond Cattell, who played a major role in the development of factor
analysis and intelligence testing also advocated eugenics, fascism, and
Nazi race science\cite{MehlerBeyondismRaymondCattell1997}. However, the
other authors named in this topic --- John Horn and Peter Bentler --- do
not appear to have contributed to race science discourse. Instead this
topic appears to identify ``mainstream'' (non-race science) intelligence
research, specifically on factor analysis and the debate over whether
intelligence in unidimensional or multidimensional. (Three other topics
only associated with mainstream intelligence journals and terms were
also identified by this topic model.) This topic indicates that the
model is not simply lumping race science research together with other
intelligence research.

The Silge plot for topic 24, by contrast, suggests a topic focused
squarely on race and intelligence, with multiple racial terms
(\texttt{whites}, \texttt{blacks}, \texttt{racial\_differences},
\texttt{races}, \texttt{race\_differences}) and the names of three
prominent Pioneer-funded researchers, \texttt{jensen}, \texttt{lynn},
and \texttt{rushton}. To confirm this impression from the Silge plot, we
independently coded the top 121 papers in this topic (those with
\(gamma > 0.97\)) as contributing to race science discourse or not. 105
(87\%) were coded as race science discourse by both authors; 10 (8.3\%)
were coded as not race science discourse by both; and discordant codes
were given to the remaining 6 (5.0\%) (95\% agreement, Cohen's
\(\kappa = 0.74\)).

In almost all years, most papers in topic 24 were published in
mainstream journals rather than MQ. Jensen's ``How Much Can We Boost IQ
and Scholastic Achievement?'' \cite{JensenHowMuchCan1969} was published
in 1969 in Harvard Educational Review (not included in this study), and
the time series indicates that, around this time, there was an increase
in articles in topic 24 in both PS and MQ (the only two journals in our
corpus that were active at the time). MQ shows another sharp increase in
the late 1970s, around the time Lynn was brought on as assistant editor,
\emph{{[}confirming the claim that Lynn shifted MQ's focus to
intelligence research{]}}. PID published multiple papers in topic 24
almost immediately after it was founded, with I showing a more gradual
increase between the mid-1970s and mid-1990s.

Notably, BG published very few articles in this topic. Outside of the
field itself, behavior genetics is strongly associated with race science
by both other academics and the general public. Panofsky
\cite{PanofskyMisbehavingScienceControversy2014} argues that, prior to
Jensen's 1969 paper, behavior genetics emphasized the study of non-human
animals and intentionally avoided associations with eugenics and public
controversy more generally. In response to Jensen, critics such as
Lewontin offered broad critiques of behavior genetics as such, and
behavior geneticists in turn adopted a radical conception of academic
freedom (without any sense of responsibility for the social implications
of academic research) and a siege or wartime mentality, as demonstrated
by Sandra Scarr's 1986 presidential address to the Behavior Genetics
Association (BGA) \cite{ScarrThreeCheersBehavior1987}. Scarr's address
coincided with the period between 1970-1990 when BG published articles
in topic 24 (including Scarr's address itself).

However, Glayde Whitney's 1995 BGA presidential address was published in
MQ rather than BG. According to Whitney, BG's editor refused to publish
the address and the BGA Executive Committee (except for Whitney himself)
voted to ``issue an official statement of denouncement'' of the address;
though we were unable to find any such a statement, and Panofsky reports
that ``nothing official was done to Whitney.'' In his address, Whitney
criticized ``the Marxist-Lysenkoist denial of genetics''; quoted Peter
Brimelow --- a white nationalist and later founder of the racist and
xenophobic website VDARE --- defining ``racist'' as ``anyone who is
winning an argument with a liberal''; and proposed as a ``reasonable
scientific hypothesis'' that differences in murder rates between
countries and cities was caused by racial genetic differences in
intelligence, empathy, aggression, and impulsivity.

The topic model analysis and the case of Whitney's address suggest that,
by the 1990s, behavior genetics had at least somewhat distanced itself
from race science.
\emph{{[}\url{https://link.springer.com/search?use-oscar-shared-search=true\&sortBy=newestFirst\&dateTo=2010\&sortOrder=newestFirst\&facet-journal-id=10519\&content-type=Article\&query=race+intelligence\&date=custom\&dateFrom=\&date-facet-mode=between\&facet-start-year=1972\&previous-start-year=1972\&facet-end-year=2010\&previous-end-year=2023}{]}}

A disciplinary history of behavior genetics published in 2010
\cite{McGueEndBehavioralGenetics2010} acknowledges the historical
association with eugenics, but does not mention Jensen and other race
science controversies since World War II. Instead postwar behavior
genetics is portrayed as an endeavor to counterbalance the hegemony of
the ``Blank Slate,'' the supposedly widespread view that genetic
differences play no role in explaining individual differences
\cite{BatesonCorpseWearisomeDebate2002}.

These findings for \emph{Behavior Genetics} complicate at least the
popular association of behavior genetics with race science, if not
Panofsky's more nuanced analysis.

\hypertarget{discussion}{%
\section*{Discussion}\label{discussion}}

Our findings indicate that \emph{Mankind Quarterly} (MQ) was less
important as an echo chamber for late 20th century race science than
mainstream psychology journals, especially \emph{Intelligence} (I) and
\emph{Personality and Individual Differences} (PID). Indeed, MQ
published a very different kind of race science from the
race-and-intelligence research published in I and PID, even after
\emph{{[}Lynn joined the MQ editorial board{]}}.

We stress that we are not identifying any particular scientist as a
racist based on the topic model results. Scientific research can be
utilized to promote scientific racism --- and thus count as race science
--- independently of the views and attitudes of the scientist who
conducted that
research.\cite{TaberyWhyStudyingGenetics2015, CarlsonQuantifyingContextualizingImpact2020, HennWhyDNANo2021}
But this does not mean that scientists are not responsible for the
downstream social effects of their research, even when these effects
involve deliberate misinterpretation.\emph{{[}Douglas, Kitcher, Kourany,
Havstad; Block and Dworkin?{]}} Like other citizens, scientists are
responsible for mitigating the reasonably foreseen harms that result
from their actions. And we now have decades of empirical evidence that
research on the genetics of intelligence --- even research that does not
itself touch on race --- can and will be used to promote scientific
racism.\cite{MeyerWrestlingSocialBehavioral2023}

\emph{{[}engagement dilemma: technical critiques can legitimize{]}}

\emph{{[}this probably goes too far even for a discussion section{]}}

Proponents of a strong conception of academic freedom often point to
chapter 2 of philosopher John Stuart Mill's \emph{On Liberty}
\cite{MillLiberty1862}, ``Of the Liberty of Thought and Discussion.''
However, in chapter 1 of the same book, Mill states the Harm Principle
as the overall thesis of his book: ``The only end for which people are
entitled, individually or collectively, to interfere with the liberty of
action of any of their number is self-protection. The only purpose for
which power can be rightfully exercised over any member of a civilized
community, against his will, is to prevent harm to others.'' Even for
Mill, free speech and free inquiry are appropriately restrained when the
ideas in question will harm others.

More recently, philosopher Shannon Dea
\cite{DeaEvolvingSocialPurpose2021} has argued that academic freedom is
importantly distinct from freedom of speech. Whereas freedom of speech
is universal, enjoyed by all residents of countries where speech is
legally protected, academic freedom is more exclusive, enjoyed only by
college and university faculty. (Though, as Dea points out,
non-tenure-track faculty often do not have effective protections for
academic freedom.) Freedom of speech primarily serves to protect the
interests of the individual speaker: that you will not be punished by
the state for stating unpopular opinions. But academic freedom is
``other-directed,'' primarily serving to benefit the common good of
society as a whole by enabling the pursuit of truth.

Advocates of research on genetics and intelligence --- both anti-racist
liberals and race scientists --- claim this work will lead to social
benefits. This reflects a recognition that the purpose of academic
freedom is to promote the common good, rather than merely satisfy an
individual's curiosity. But the promised benefits are either vague or do
not actually require understanding the genetics of
intelligence.\cite{KaplanGaltonQuincunxProbabilistic2021, CoopLotteryLuckLegacy2022}

Using agent-based models, Weatherall et
al.\cite{WeatherallHowBeatScience2018} show that a strategy of
``selective sharing'' --- promoting favorable results from independent
research --- can be effective at misleading a target audience.

\emph{{[}for discussion: the possibility for white supremacist
appropriation, ie, race science done by scientists who are not
themselves scientific racists:
CarlsonQuantifyingContextualizingImpact2020; probably some reviews of
Paige Harden; O'Connor and Weatherall{]}}

### 2023-06-29 paper/supplement_paper.tex
working on paper

\hypertarget{data-preparation}{%
\subsection*{Data preparation}\label{data-preparation}}

@@

\hypertarget{vocabulary-selection}{%
\subsubsection*{Vocabulary selection}\label{vocabulary-selection}}

@@

We also adapted the ``discursive space'' analysis proposed by
\emph{{[}Hicks 2021?{]}}. This analysis combines the Hellinger distance
between topic-document distributions --- capturing how ``far apart'' two
documents are --- with dimensional reduction techniques used for data
visualization. This produces a two-dimensional visualization of
individual documents, arranged so that documents with similar topic
distributions are visually clustered and visually distinct from
documents with different topic distributions. \emph{{[}availability{]}}

@@

>>>>

\hypertarget{text-preparation}{%
\subsection*{Text preparation}\label{text-preparation}}

@@

\hypertarget{data-and-code-availability}{%
\subsubsection*{Data and code
availability}\label{data-and-code-availability}}

Document metadata retrieved from Crossref does not appear to be covered
by any copyright or other intellectual property restrictions. However,
due to copyright restrictions, we are unable to make document fulltext
or Web of Science search results publicly available. Code for the corpus
assembly and text preparation steps described above is available by
request.

The public analysis repository,
\url{https://github.com/dhicks/race-science} \emph{{[}DOI{]}}, includes
document metadata and documentwise counts of NLP-extracted noun phrases
(``document-term matrices''), along with code to reproduce the analysis
of the following sections. \emph{{[}repo readme: requires R 4.1.2; rig
(\url{https://github.com/r-lib/rig/}) for managing multiple R
installations{]}}

\hypertarget{vocabulary-selection}{%
\subsection*{Vocabulary selection}\label{vocabulary-selection}}

@@

\emph{{[}drop{]}} We also adapted the ``discursive space'' analysis
proposed by \emph{{[}Hicks 2021?{]}}. This analysis combines the
Hellinger distance between topic-document distributions --- capturing
how ``far apart'' two documents are --- with dimensional reduction
techniques used for data visualization. This produces a two-dimensional
visualization of individual documents, arranged so that documents with
similar topic distributions are visually clustered and visually distinct
from documents with different topic distributions.
\emph{{[}availability{]}}

@@

\emph{{[}rewrite{]}}
